### 2018-05-13 src/tex/util/variables.tex

%------------------------------------------
% variables.tex
% All key-value pair should be set here
%------------------------------------------

% Variables declaration %
\def \judul {Pengembangan Algoritma Pencarian Non Interaktif untuk Penyelesaian Permasalahan Pencarian Ulam dengan Kebohongan Jamak}
\def \penulis {Risyanggi Azmi Faizin}
\def \oj {Sphere Online Judge}
\def \soal {Guess The Number v5}
\def \nomorsoal {1598}
\def \nrplama {5116201052}
\def \nrpbaru {05111650010052}
\def \jurusanlama {Jurusan Teknik Informatika}
\def \jurusanbaru {Departemen Informatika}
\def \fakultaslama {Fakultas Teknologi Informasi}
\def \fakultasbaru {Fakultas Teknologi Informasi dan Komunikasi}
\def \pembimbingsatu {Dr. Ir. R. V. Hari Ginardi, M.Sc}
\def \nikpembimbingsatu {196505181992031003}

\def \juduleng {Development of Non-Interactive Searching Algorithm for Solving Ulam's Searching Problem with Many Lies}
\def \jurusanlamaeng {Informatics Department}
\def \jurusanbarueng {Informatics Department}
\def \fakultaslamaeng {Faculty of Information Technology}
\def \fakultasbarueng {Faculty of Information and Communication Technology}

\definecolor{abucover}{RGB}{128,128,128}